---
title: "e19.10_dilution_curve_prep_qc"
author: "Chandler Sutherland"
date: "2025-01-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
require(tidyverse)
library(openxlsx)
library(patchwork)
```

Copyright © Chandler Sutherland Email: [chandlersutherland\@berkeley.edu](mailto:chandlersutherland@berkeley.edu){.email}

Purpose: produce a custom QC report for the E19.10 experiment. 

All of these libraries are based on the same gDNA source: Col-0 planting ID 1 plant ID 8. Libraries 19-24 are based on the same pre-PCR library following the nanoseq vCAS protocol, while Library 25 is from a different pre-PCR library following the nanoseq protocol as published. Libraries 19-24 were diluted to different magnitudes before PCR amplification. 

# CS19-24 (dilutions)

Beginning with nanoseq vCAS: 

- performed a step 0 ampure bead cleanup (1x) to remove the AE buffer I decided to elute in (don't do again)
- of the 4 gDNA samples (plants 7-10), I added H2O to get 4 2.5 ng/ul samples in 250ul. 
- eluted in 25ul, 20ul for nanoseq vCAS and 5ul for QC

## Pre-pcr quantification and library size 

```{r}
cleanup_1 <- data.frame(
  plant_id = c(7:10), 
  quibit_conc = c(16.1, 10.2, 17.0, 16.7)) %>%
  mutate(ng_yield = quibit_conc*20) %>% 
  mutate(percent_recover=ng_yield/625*100)

cleanup_1
```

recovered 30-53% of gDNA from original

Started the nanoseq protocolvCAS, including bead cleanup, fragmentation, A-tailing, ligation, and a two sided bead cleanup. Elute in 50ul NFW, and do qPCR, quibit, and frag analyzer. 

Frag analyzer submission #: M005770_CS 
qPCR result: 20250107 library check
qPCR method: SYBR qPCR 2 (quibit estimate of libraries)

```{r}
lib_pt1 <- data.frame(
  plant_id = c(7:10),
  quibit_conc = c(2.86, 1.49, 3.23, 2.01), #ng/ul
  avg_frag_size = c(395, 399, 466, 407), #from smear analysis 
  frag_conc = c(1.26, 0.66, 1.24, 0.95), #from smear analysis 
  qpcr_sybr_2 = c(0.213, 0.165, 0.155, 0.104) #nM 
  ) %>%
  mutate(fmol_yield=qpcr_sybr_2*50) %>%
  mutate(qpcr_ng_ul=qpcr_sybr_2/10^6*660*avg_frag_size) %>%
  mutate(ng_yield=qpcr_ng_ul*50)

yield_1 <- merge(cleanup_1, lib_pt1, by = 'plant_id') %>% 
  mutate(percent_ng_yield=ng_yield.y/ng_yield.x*100) %>% 
  subset(select=c(plant_id, avg_frag_size, qpcr_sybr_2, fmol_yield, percent_ng_yield))

yield_1
```
Relative to the amount of DNA input into the first part of the pipeline, only recover 0.4-1%. However, this is plenty of fmol to work with before PCR amplification. 


Look at the fragment distribution: 
```{r}
#load data 
frag_traces_1 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250114\\2025 01 10 17H 40M Electropherogram (1).csv") %>%
  pivot_longer(cols=starts_with('F'), names_to='sample', values_to='Intensity') %>% 
  separate('sample', c(NA, 'sample'))  %>% 
  rename(size = `Size (bp)`) %>% 
  mutate(plant_id=case_when(sample=='CS14' ~ 7, 
                            sample=='CS15' ~ 8,
                            sample=='CS16' ~ 9, 
                            sample=='CS17' ~ 10))

frag_traces_1$plant_id <- as.factor(frag_traces_1$plant_id)

pre_pcr <- ggplot() +
  geom_line(data=frag_traces_1, aes(x=size, y= Intensity, color=plant_id))+
  scale_x_log10(limits=c(5,NA), breaks=c(1, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')

pre_pcr
```
8 and 10 have the best curves, 8 has more fmol. Proceed with this prep for sequencing. 

## Dilution and re-qPCR 

The goal of this experiment is a dilution curve for eventual sequencing. So dilute the 8 pre-PCR library above across a range of values. Then qPCR check with SYBR. Attempted to KAPA check, only got two samples. Estimate the rest with KAPA conversion factor (see qpcr_troubleshooting.Rmd)

```{r}
lib8_nM <- 0.165	#used the smear analysis frag size instead of smooth. This is what is sequenceable. So add nM_estimated, which is 

dilution <- data.frame(
  lib_id = c('CS_19', 'CS_20', 'CS_21', 'CS_22', 'CS_23', 'CS_24'), 
  fmol_input_ideal = c(2, 1, 0.5, 0.25, 0.125, 0.0625), 
  vol_added = c(12.11, 4.84, 2.42, 1.21, 0.60, 0.30)) %>%
  mutate(nM_ideal = fmol_input_ideal/20)
```

SYBR_qPCR results: 20250121 dilution curve quant 1(SYBR)
KAPA_qPCR results: 20250122 dilution curve quant 2.1(KAPA)
```{r}
kapa_factor <- 2.56887
measured_dilution <- dilution %>% 
  mutate(sybr_qpcr_2 = c(0.0818, 0.0452, 0.0281, 0.0101, 0.0062, 0.0039)) %>%
  mutate(kapa_measured = c(NA, 0.0672, 0.0315, NA, NA, 0.0049)) %>%
  mutate(kapa_estimated=sybr_qpcr_2*kapa_factor)

#apples to apples SYBR comparison 
dilution_sybr_nM <- measured_dilution %>% subset(select=c(lib_id, nM_ideal, sybr_qpcr_2)) %>%
  mutate(percent_dif = (nM_ideal-sybr_qpcr_2)/nM_ideal*100)

dilution_sybr_nM

dilution_fmol <- measured_dilution %>% 
  mutate(sybr_fmol = sybr_qpcr_2*20, kapa_fmol=kapa_estimated*20) %>% 
  subset(select=c(lib_id, fmol_input_ideal, sybr_fmol, kapa_fmol))
```

Pretty close to the dilution values I had hoped. This particular qPCR also had high variation between replicates, so may be ok to go with nM ideal. 

## Post-PCR quantification and fragment distribution

SYBR qPCR: 20250123 library quant sybr and nano AND 20250128 library quant sybr, see 19.10 calibration benchling for more details 

```{r}
post_pcr <- dilution %>% subset(select=c(lib_id, fmol_input_ideal, nM_ideal)) %>%
  mutate(sybr_qpcr_2 = c(32.513, 11.115, 19.339, 11.459, 2.050, 3.870)) %>%
  mutate(avg_frag_size=c(523, 545, 518, 483, 582, 392)) %>% 
  mutate(frag_conc = c(9.9645, 3.6445, 7.7771, 3.7213, 1.0226, 0.7635)) %>%
  mutate(quibit_conc = c(13.3, 6.44, 11.6, 5.42, 1.73, 1.13))

post_pcr

pcr_effi <- post_pcr %>% mutate(pcr_cycles = c(11:16)) %>%
  mutate(fmol_output_ideal = fmol_input_ideal*2^pcr_cycles) %>%
  mutate(fmol_output_sybr = sybr_qpcr_2*50) %>%
  mutate(pcr_yield = fmol_output_sybr/fmol_output_ideal*100) %>% subset(select=c(lib_id, 
                                                                                 fmol_output_ideal,
                                                                                 fmol_output_sybr,
                                                                                 pcr_cycles,
                                                                                 pcr_yield))

pcr_effi
```

There's more I need to do here (or could do here) with calculating PCR efficiency/cleanup efficiency and check what happened with CS_23/24 

Fragments! 
fragment size distribution: 20250124 pilot v 2 post pcr/M005825_CS
```{r}
#load data 
frag_traces_2 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250124 pilot v 2 post pcr\\2025 01 22 16H 57M Electropherogram.csv") %>%
  pivot_longer(cols=starts_with('E'), names_to='sample', values_to='Intensity') %>% 
  separate('sample', c(NA, 'sample'))  %>% 
  rename(size = `Size (bp)`) %>% 
  mutate(sample=case_when(sample=='CS19' ~ 'CS_19', 
                            sample=='CS20' ~ 'CS_20',
                            sample=='CS21' ~ 'CS_21', 
                            sample=='CS22' ~ 'CS_22', 
                            sample=='CS23' ~ 'CS_23', 
                            sample=='CS24' ~ 'CS_24'))

frag_traces_2$sample <- as.factor(frag_traces_2$sample)

#plot!
post_pcr_plot <- ggplot() +
  geom_line(data=frag_traces_2, aes(x=size, y= Intensity, color=sample))+
  scale_x_log10(limits=c(5,NA), breaks=c(1, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')

post_pcr_plot
```

The plots look similar to the distributions reported by novogene for the finished libraries: same multi-modal peaks across libraries, same-ish distribution. I think CS_23 and 24 if they were more concentrated would look the same. Overall OK, and similar enough for sequencing that I should be alright. 


# CS25 (control)

CS25 was prepped using the standard library protocol. Started with 12.5ul of Col-0 8 gDNA + 7.5ul H2O (50ng). Did not do fragment analysis at the pre-pcr stage. 

## Pre-pcr quantification
```{r}
lib_pt2 <- data.frame(
  rep_id = c(1, 2),
  #quibit_conc = c(NA, NA), #ng/ul
  #avg_frag_size = c(NA, NA), #from smear analysis 
  #frag_conc = c(NA, NA), #from smear analysis 
  qpcr_sybr_2 = c(0.1998, 0.0843), #nM 
  qpcr_kapa = c(0.196, 0.1519)
  ) %>%
  mutate(fmol_yield=qpcr_sybr_2*50) %>%
  mutate(qpcr_ng_ul=qpcr_sybr_2/10^6*660*573) %>% #use average fragment size at this step 
  mutate(ng_yield=qpcr_ng_ul*50) %>% 
  mutate(percent_ng_yield=ng_yield/50*100) %>%
  mutate(fmol_per_ng=fmol_yield/50)

lib_pt2
```

Recovered 3 and 7% of the input ng. This is less than my previous library prep using standard nano protocol (10-30%), and the reported recovery is higher. The fmol/input ng reported in the bottleneck optimization document is 0.35. Within previous runs though (0.1-0.34 fmol/ng). Higher than vCAS, which has ~1% recovery, but not by that much. 

## Post-PCR quantification and fragment analysis 

Moved forward with rep 2. Since we are following nanoseq as published, use the KAPA value. The presumed ideal is 0.5fmol (KAPA)/0.194 fmol (SYBR). Took 2.88ul of rep 2 for 25ul (extra few ul for qPCR in the future should I so choose). 20ul forward to PCR, 15 cycles. 

```{r}
post_pcr_2 <- data.frame(
  lib_id = c('CS_25'), 
  fmol_input_ideal = c(0.194), 
  sybr_qpcr_2 = c(5.163), 
  avg_frag_size=c(692), 
  frag_conc=c(3.391),
  quibit_conc=c(6.37)) %>%
  mutate(nM_ideal=fmol_input_ideal/20)

pcr_effi_2 <- post_pcr_2 %>% mutate(pcr_cycles = c(15)) %>%
  mutate(fmol_output_ideal = fmol_input_ideal*2^pcr_cycles) %>%
  mutate(fmol_output_sybr = sybr_qpcr_2*50) %>%
  mutate(pcr_yield = fmol_output_sybr/fmol_output_ideal*100) %>% subset(select=c(lib_id, 
                                                                                 fmol_output_ideal,
                                                                                 fmol_output_sybr,
                                                                                 pcr_cycles,
                                                                                 pcr_yield))

pcr_effi_2
```
Slightly higher fmol output ideal, but this input is between cycle recommendations. Similar % yield as the other low input libraries. PCR is just more efficient at higher concentrations I think. 

Fragment!
```{r}
#load data 
frag_traces_25 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250124 pilot v 2 post pcr\\2025 01 24 17H 38M Electropherogram.csv") %>%
  rename(size = `Size (bp)`, Intensity=`B10: CS_25`) %>% 
  mutate(sample='CS_25')

frag_traces_25$sample <- as.factor(frag_traces_25$sample)

#plot!
post_pcr_plot2 <- ggplot() +
  geom_line(data=frag_traces_25, aes(x=size, y= Intensity, color=sample))+
  scale_x_log10(limits=c(5,NA), breaks=c(1, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')

post_pcr_plot2
```

# Summary 

Nanoseq vCAS recovers 0.5-1% of input gDNA, with improved fragment analyzer peaks compared to 7% in Nanoseq as is. 

