---
title: "E19.12.2_qc_lib_prep_final"
author: "Chandler Sutherland"
date: "2025-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
require(tidyverse)
library(openxlsx)
library(patchwork)
library(ggbeeswarm)
```

Copyright © Chandler Sutherland Email: [chandlersutherland\@berkeley.edu](mailto:chandlersutherland@berkeley.edu){.email}

Purpose: produce a custom QC report for the E19.12 experiment. 

## Yield 

Two components on yield: the pre PCR input and the post PCR. This will inform how many cycles to do next time and if my input is appropriate. 14 cycles this time. Started with 500ng of each gDNA sample. An extra ampure step at the beginning so that volume is consistent going in to the protocol. 

```{r}
#quibit estimates of yield 
quibit_yield <- data.frame(library=c('library 1', 'library 2', 'library 3', 'library 4', 'library 5', 'library 6', 'library 7', 'library 8', 'library 9'), 
           quibit_pre_pcr=c(1.01, 0.787, 0.887, 0.767, 0.652, 0.640, 0.740, 0.527, 0.411), 
           quibit_post_pcr=c(6.48, 3.75, 3.03, 5.09, 6.04, 3.47, 7.27, 5.59, 4.46)) %>% 
  mutate(input=500) %>%
  mutate(pre_pcr_yield=quibit_pre_pcr*100)%>%
  mutate(post_pcr_yield=quibit_post_pcr*50)

q_yield <-  quibit_yield %>% 
  subset(select=c(library, input, pre_pcr_yield, post_pcr_yield)) %>% 
  pivot_longer(cols=!library, names_to='timepoint', values_to='ng')
  
q_yield$timepoint <- factor(q_yield$timepoint, levels=c('input', 'pre_pcr_yield', 'post_pcr_yield'))

ggplot(q_yield, aes(x=timepoint, y=ng, color=library, group=library))+
  geom_point()+
  geom_line()+
  ylab('ng yield')+
  theme_classic()

quibit_yield %>% mutate(pre_pcr_percent=quibit_pre_pcr/input*100) %>% subset(select=c(library, pre_pcr_percent))

metric_table <- quibit_yield %>% subset(select=c(library, quibit_pre_pcr, quibit_post_pcr))
```

Recovered 0.08-0.17% of the input ng before PCR. Less than last library prep (0.4-1%), but I also started with more that time, and maybe % yield is dependent on input. Still plenty of library for amplification. 

## Fragment Distribution 

### Pre-pcr 
Pre-pcr:M006217_CS

I did normalize by concentration before submitting, so that was less of a factor in the trace estimates of average fragment. In general, I think more concentrated is better. 

Col-0 gDNA 7	CS35
Col-0 gDNA 9 	CS36
Col-0 gDNA 10 	CS37
msh2 gDNA 2	CS38
msh2 gDNA 4	CS39
msh2 gDNA 7	CS40
uni1d gDNA 1	CS41
uni1d gDNA 2	CS42
uni1d gDNA 3	CS43

```{r}
pre_pcr_smear <- data.frame(library=c('library 1', 'library 2', 'library 3', 'library 4', 'library 5', 'library 6', 'library 7', 'library 8', 'library 9'), 
                       avg_frag = c(474, 405, 463, 412, 401, 428, 475, 389, 383), 
                       percent_200_1000 = c(76.2, 69.9, 77.7, 71.3, 68.2, 73.9, 76.3, 65.0, 60.5))

pre_pcr_smear %>% pull(avg_frag) %>% mean()

metric_table <- pre_pcr_smear %>% 
  subset(select=-percent_200_1000) %>% 
  rename('frag_pre_pcr'='avg_frag') %>% 
  merge(metric_table, on=library)
```

```{r}
#load data to plot 
run_1 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250603 pre pcr 2\\2025 06 02 15H 25M Electropherogram.csv") %>%
  pivot_longer(cols=c(`H5: CS39`,`H6: CS43`), names_to='sample', values_to='Intensity') 

run_2 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250603 pre pcr 2\\2025 06 02 14H 20M Electropherogram.csv") %>%
  pivot_longer(cols=c(`G9: CS35`,`G11: CS42`), names_to='sample', values_to='Intensity') 

run_3 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250603 pre pcr 2\\2025 06 02 11H 53M Electropherogram.csv") %>%
  pivot_longer(cols=`E6: CS36`:`E11: CS41`, names_to='sample', values_to='Intensity') 


frag_traces_1 <- rbind(run_1, run_2, run_3) %>% 
  separate('sample', c(NA, 'sample_number')) %>% 
  rename(size = `Size (bp)`) %>% 
  mutate(Library=case_when(sample_number=='CS35' ~ 'Library 1', 
                            sample_number=='CS36' ~ 'Library 2',
                            sample_number=='CS37' ~ 'Library 3', 
                            sample_number=='CS38' ~ 'Library 4', 
                            sample_number=='CS39' ~ 'Library 5', 
                            sample_number=='CS40' ~ 'Library 6', 
                            sample_number=='CS41' ~ 'Library 7', 
                            sample_number=='CS42' ~ 'Library 8', 
                            sample_number=='CS43' ~ 'Library 9'
                            ))%>%
  mutate(plant_id=case_when(sample_number=='CS35' ~ 'Col0_7', 
                            sample_number=='CS36' ~ 'Col0_9',
                            sample_number=='CS37' ~ 'Col0_10', 
                            sample_number=='CS38' ~ 'msh2_2', 
                            sample_number=='CS39' ~ 'msh2_4', 
                            sample_number=='CS40' ~ 'msh2_7', 
                            sample_number=='CS41' ~ 'uni1d_1', 
                            sample_number=='CS42' ~ 'uni1d_2', 
                            sample_number=='CS43' ~ 'uni1d_3'
                            )) %>%
  separate(plant_id, sep='_', into=c('genotype', NA), remove=FALSE)

frag_traces_1$plant_id <- as.factor(frag_traces_1$plant_id)

pre_pcr_2 <- ggplot() +
  geom_line(data=frag_traces_1, aes(x=size, y= Intensity, color=plant_id))+
  scale_x_log10(limits=c(10,NA), breaks=c(10, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  ylim(0, 2000)+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')+
  facet_wrap(~genotype)

pre_pcr_2
```
### Post-pcr
Post PCR: M006241_CS

I did normalize by concentration before submitting, so that was less of a factor in the trace estimates of average fragment. In general, I think more concentrated is better. 

Col-0 gDNA 7	CS44
Col-0 gDNA 9 	CS45
Col-0 gDNA 10 	CS46
msh2 gDNA 2	CS47
msh2 gDNA 4	CS48
msh2 gDNA 7	CS49
uni1d gDNA 1	CS50
uni1d gDNA 2	CS51
uni1d gDNA 3	CS52

```{r}
post_pcr_smear <- data.frame(library=c('library 1', 'library 2', 'library 3', 'library 4', 'library 5', 'library 6', 'library 7', 'library 8', 'library 9'), 
                       avg_frag = c(594, 515, 588, 537, 531, 585, 582, 506, 577), 
                       percent_200_1000 = c(92.8, 97.3, 94.8, 97.5, 97.2, 95.3, 92.9, 98.4, 94.9))

post_pcr_smear %>% pull(avg_frag) %>% mean()

metric_table <- post_pcr_smear %>% 
  subset(select=-percent_200_1000) %>% 
  rename('frag_post_pcr'='avg_frag') %>% 
  merge(metric_table, on=library)
```

Larger average fragment size, but almost everything within 200-1000. 

```{r}
#load data to plot 
run_4 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250604 post pcr\\2025 06 04 19H 45M Electropherogram.csv") %>%
  pivot_longer(cols=`E1: CS44`:`E9: CS52`, names_to='sample', values_to='Intensity') 

frag_traces_2 <- run_4 %>% 
  separate('sample', c(NA, 'sample_number')) %>% 
  rename(size = `Size (bp)`) %>% 
  mutate(Library=case_when(sample_number=='CS44' ~ 'Library 1', 
                            sample_number=='CS45' ~ 'Library 2',
                            sample_number=='CS46' ~ 'Library 3', 
                            sample_number=='CS47' ~ 'Library 4', 
                            sample_number=='CS48' ~ 'Library 5', 
                            sample_number=='CS49' ~ 'Library 6', 
                            sample_number=='CS50' ~ 'Library 7', 
                            sample_number=='CS51' ~ 'Library 8', 
                            sample_number=='CS52' ~ 'Library 9'
                            ))%>%
  mutate(plant_id=case_when(sample_number=='CS44' ~ 'Col0_7', 
                            sample_number=='CS45' ~ 'Col0_9',
                            sample_number=='CS46' ~ 'Col0_10', 
                            sample_number=='CS47' ~ 'msh2_2', 
                            sample_number=='CS48' ~ 'msh2_4', 
                            sample_number=='CS49' ~ 'msh2_7', 
                            sample_number=='CS50' ~ 'uni1d_1', 
                            sample_number=='CS51' ~ 'uni1d_2', 
                            sample_number=='CS52' ~ 'uni1d_3'
                            )) %>%
  separate(plant_id, sep='_', into=c('genotype', NA), remove=FALSE)

frag_traces_2$plant_id <- as.factor(frag_traces_2$plant_id)

post_pcr_2 <- ggplot() +
  geom_line(data=frag_traces_2, aes(x=size, y= Intensity, color=plant_id))+
  scale_x_log10(limits=c(10,NA), breaks=c(10, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  ylim(0, 2000)+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')+
  facet_wrap(~genotype)

post_pcr_2
```

```{r}
frag_traces_1 %>% mutate(pcr='pre') %>% rbind(frag_traces_2 %>% mutate(pcr='post')) %>%
  ggplot() +
  geom_line(aes(x=size, y= Intensity, color=pcr))+
  scale_x_log10(limits=c(10,NA), breaks=c(10, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  ylim(0, 2000)+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')+
  facet_wrap(~plant_id)
```

```{r}
size_dif <- pre_pcr_smear %>% mutate(pcr='pre') %>%
  rbind(post_pcr_smear %>% mutate(pcr='post')) %>%
  subset(select=-percent_200_1000)%>%
  pivot_wider(names_from=pcr, values_from=avg_frag) %>%
  mutate(dif=post-pre)

size_dif %>% pull(post) %>% mean()
```
Expect primers to add 2*37 bp to each fragment. So there must be some enrichment of larger fragments that get amplified. Or the smaller parts of my trace are not amplifiable, or less accurate estimate at lower concentration. 

### Final 
Final pooled fragment distribution: M006248_CS

```{r}
frag_traces_3 <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\fragment analysis\\20250610 pool\\2025 06 09 17H 17M Electropherogram.csv") %>% 
  rename(size = `Size (bp)`, KVKCS002 = `F2: KVKCS002`)

frag_traces_3 %>% ggplot() +
  geom_line(aes(x=size, y= KVKCS002))+
  scale_x_log10(limits=c(10,NA), breaks=c(10, 100, 200, 300, 400,500, 600, 700, 800, 1000, 1500,3000,6000))+
  ylim(0, 2000)+
  theme_classic()+
  geom_vline(xintercept=200, linetype=2, color='grey')+
  geom_vline(xintercept=1000, linetype=2, color='grey')+
  annotate('text', x=558, y=2000, label='558')+
  xlab('Size (bp)')+
  ylab('Intensity')+
  labs(title='KVKCS002 pool fragment distribution')
```

The average of the post-PCR smear analysis was 557, so this is right on the money. Cool!

## qPCR concentration 

3 measurements: initial estimate of the library, estimate of the dilution, and post-PCR quantification before pooling. Changes to analysis: set baseline manually: 1-3 cycles, then Ct threshold to 0.2-0.5 depending on the run.

### Pre-PCR
Initial library estimate: 2025052. This is the updated, standardized estimates (not necessarily what I thought the first time around). For that, see the other E19.12 qc library prep document 

```{r}
pre_pcr_q <- read.xlsx("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\Library qPCR\\20250527 - frag update - 0610 recalc.xlsx", sheet='All dilution final', startRow=2, cols=c(1:3, 6)) %>%
  rename('library' = 'X1', 'replicate' = 'X2', 'gblock_estimate' = 'gblock.nM', 'benchmark_estimate'='X4') %>%
  mutate(library=str_to_lower(library))

pre_pcr_q
```


```{r}
pos <- read_csv('pre_pcr_qPCR_positive_controls.csv') %>% 
  mutate(method='gblock_estimate') 

pos_control <- pre_pcr_q %>% filter(str_detect(library, '^positive')) %>% 
  mutate(date='20250527_recalc') %>%
  pivot_longer(cols=c('gblock_estimate', 'benchmark_estimate'), names_to='method', values_to='estimate') %>%
  rbind(pos)

ggplot(pos_control)+
  geom_boxplot(aes(x=library, y=estimate))+
  geom_point(aes(x=library, y=estimate, color=date, shape=replicate), position='jitter')+
  scale_color_manual(values=c("20250527_recalc" = 'purple'))+
  #scale_y_log10()+
  theme_classic()
```
```{r}
comparison <- pre_pcr_q %>%
  pivot_longer(cols=c('gblock_estimate', 'benchmark_estimate'), names_to='method', values_to='estimate')
  

ggplot(comparison)+
  geom_boxplot(aes(x=library, y=estimate, color=method), position=position_dodge2(width=0.5))+
#  geom_point(aes(x=library, y=estimate, color=method), position=position_dodge2(width=0.5))+
  theme_classic()

pre_pcr_q %>%
  ggplot()+
  geom_point(aes(x=gblock_estimate, y=benchmark_estimate, color=library))+
  geom_abline(intercept=0, slope=1)+
  theme_classic()
```


```{r}
lib_pos <- pre_pcr_q %>% filter(!str_detect(library, '^benchmark')) %>% subset(select=c(library, replicate, gblock_estimate))

ggplot(lib_pos)+
  geom_boxplot(aes(x=library, y=gblock_estimate))+
  geom_point(aes(x=library, y=gblock_estimate, color=replicate))+
  scale_color_manual(values=c('dilution 1' = 'lightblue', 
                              'dilution 2' = 'lightblue', 
                              'dilution 3' = 'lightblue', 
                              'dilution 4' = 'darkblue', 
                              'dilution 5' = 'darkblue', 
                              'dilution 6' = 'darkblue'))+
  ylim(0, 0.3)+
  ylab('nM library estimate')+
  xlab('')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))



```
Dilution 1, 2, 3 are independent 1:10 dilutions, dilution 4, 5, 6 are independent 1:100 dilutions. Not huge difference in estimates! 

```{r}
lib_only <- pre_pcr_q %>% filter(str_detect(library, '^library')) %>% subset(select=c(library, replicate, gblock_estimate))

est_pre_pcr <- lib_only %>% group_by(library) %>% 
  summarize(estimate=mean(gblock_estimate), sd=sd(gblock_estimate))%>%
  mutate(y_min = estimate-sd) %>%
  mutate(y_max=estimate + sd)

ggplot() +
  geom_bar(est_pre_pcr, mapping=aes(x=library, y=estimate), stat='identity', alpha=0.5) +
  geom_errorbar(est_pre_pcr, mapping=aes(x=library, ymin=y_min, ymax=y_max), width=0.4, alpha=0.9, position='dodge') +
  geom_point(lib_pos, mapping=aes(x=library, y=gblock_estimate))+
  ggtitle("Library Estimate (confidence intervals)")+
  ylab('nM')+
  theme_classic()

metric_table <- est_pre_pcr %>% 
  subset(select=c(library, estimate)) %>% 
  rename('nM_qpcr_pre_pcr'='estimate') %>% 
  mutate(library=str_to_lower(library)) %>%
  merge(metric_table, on=library)
```


### Dilution Check  

3 tubes were attempted, one was moved forward. What was the secondary estimate of that tube? 
```{r}
dilution_q <- read.xlsx("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\Library qPCR\\20250527_2 - frag update 0610 recalc.xlsx", sheet='Final Result', startRow=2, cols=c(1:4, 7)) %>%
  rename('library' = 'X1', 'replicate' = 'X3', 'gblock_estimate' = 'gblock.nM', 'benchmark_estimate'='bench.nM') %>%
  mutate(library=str_to_lower(library))

dilution_q

pos_control_2 <- dilution_q %>% filter(str_detect(library, '^positive')) %>% 
  mutate(date='20250527_2_recalc') %>%
  pivot_longer(cols=c('gblock_estimate', 'benchmark_estimate'), names_to='method', values_to='estimate') %>%
  subset(select=-Tube)%>%
  rbind(pos)

ggplot(pos_control_2)+
  geom_boxplot(aes(x=library, y=estimate))+
  geom_point(aes(x=library, y=estimate, color=date, shape=replicate), position='jitter')+
  scale_color_manual(values=c("20250527_2_recalc" = 'purple'))+
  #scale_y_log10()+
  theme_classic()
```
```{r}
comparison_2 <- dilution_q %>%
  pivot_longer(cols=c('gblock_estimate', 'benchmark_estimate'), names_to='method', values_to='estimate')
  

ggplot(comparison_2)+
  geom_boxplot(aes(x=library, y=estimate, color=method), position=position_dodge2(width=0.5))+
#  geom_point(aes(x=library, y=estimate, color=method), position=position_dodge2(width=0.5))+
  theme_classic()

dilution_q %>%
  ggplot()+
  geom_point(aes(x=gblock_estimate, y=benchmark_estimate, color=library))+
  geom_abline(intercept=0, slope=1)+
  theme_classic()
```
```{r}
lib_dil <- dilution_q %>% 
  filter(str_detect(library, '^library')) %>% 
  subset(select=c(library, replicate, Tube, gblock_estimate, benchmark_estimate)) %>%
  filter(Tube=='tube 1') %>% 
  pivot_longer(cols=c(gblock_estimate, benchmark_estimate), names_to = 'method', values_to='estimate')

ggplot(lib_dil)+
  geom_hline(yintercept=0.01, color='grey')+
  geom_hline(yintercept=0.012, color='grey', linetype=2)+
  geom_hline(yintercept=0.008, color='grey', linetype=2)+
  geom_boxplot(aes(x=library, y=estimate, color=method))+
  geom_point(aes(x=library, y=estimate, color=method, shape=replicate), position=position_dodge(width=0.6))+
  ylim(0, 0.02)+
  ylab('nM dilution estimate')+
  xlab('')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title='nM estimate of Tube 1')
```
```{r}
df_tbl <- data.frame(library=c('library 1', 'library 2', 'library 3', 'library 4', 'library 5', 'library 6', 'library 7', 'library 8', 'library 9'), df=c(13.76, 12.07, 14.87, 13.18, 11.38, 9.11, 9.40, 10.89, 4.47))

dil_est <- dilution_q %>% 
  filter(str_detect(library, '^library')) %>% 
  subset(select=c(library, replicate, Tube, gblock_estimate, benchmark_estimate)) %>%
  pivot_longer(cols=c(gblock_estimate, benchmark_estimate), names_to = 'method', values_to='estimate') %>%
  merge(df_tbl, on=library) %>% 
  mutate(lib_est=estimate*df) %>% 
  mutate(date='20250527_2') %>%
  subset(select=c(library, replicate, method, lib_est, date))

dil_comp <- pre_pcr_q %>% filter(str_detect(library, '^library')) %>%
  pivot_longer(cols=c(gblock_estimate, benchmark_estimate), names_to = 'method', values_to='lib_est') %>%
  mutate(date='20250527') %>%
  rbind(dil_est)

dil_comp %>%
 # filter(method=='gblock_estimate') %>%
  ggplot()+
  geom_boxplot(aes(x=library, y=lib_est, color=date), position=position_dodge(width = 0.6))+
  geom_point(aes(x=library, y=lib_est, color=date), position=position_dodge(width = 0.6))+
  theme_classic()+
  ylim(0, 0.25)+
  facet_wrap(~method)
```
```{r}
dil_estimate <- lib_dil %>% 
  filter(method=='gblock_estimate') %>% 
  group_by(library) %>% 
  summarize(estimate=mean(estimate))

metric_table <- dil_estimate %>%  
   subset(select=c(library, estimate)) %>% 
   rename('nM_qpcr_dilution'='estimate') %>% 
   merge(metric_table, on=library)
```


Both are bad! Lol. Whatever. 

### Post-PCR 
```{r}
post_pcr_q_1 <- read.xlsx("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\Library qPCR\\20250604.xlsx", sheet='Final Result', startRow=2, cols=c(1:3)) %>%
  rename('library' = 'X1', 'replicate' = 'X2', 'gblock_estimate' = 'gblock.nM') %>%
  mutate(library=str_to_lower(library)) %>%
  mutate(date='20250604')

post_pcr_q_2 <- read.xlsx("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E19\\Library qPCR\\20250606.xlsx", sheet='Final Result', startRow=2, cols=c(1:3)) %>%
  rename('library' = 'X1', 'replicate' = 'X2', 'gblock_estimate' = 'nM.0606') %>%
  mutate(library=str_to_lower(library))%>%
  mutate(date='20250606')

post_pcr_q <- rbind(post_pcr_q_1, post_pcr_q_2)
```

```{r}
final_estimate <- data.frame(library=c('library 1', 'library 2', 'library 3', 'library 4', 'library 5', 'library 6', 'library 7', 'library 8', 'library 9'), 
                             nM_qpcr_post_pcr = c(10.214, 8.361, 4.495, 11.673, 15.715, 5.412, 10.908, 17.345, 7.524))

post_pcr_q %>%
  filter(str_detect(library, '^library'))%>%
  ggplot(aes(x=library, y=gblock_estimate))+
  geom_boxplot(aes(x=library, y=gblock_estimate, color=date))+
  geom_point(aes(x=library, y=gblock_estimate, color=date, shape=replicate), position=position_dodge(width=0.6))+
  geom_point(data=final_estimate, aes(x=library, y=nM_qpcr_post_pcr), color='black', size=2)+
  ylab('nM dilution estimate')+
  xlab('')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title='nM estimates post-pcr')
```
## Summary 

All of the metrics of this library: 
```{r}
metric_table <- final_estimate %>%  
   merge(metric_table, on=library)
metric_table
```

